---
title: "COVID19 - Vaccination"
author: "Sebastien Lacube - 2021"
date: "23/08/2021"
output: html_document
bibliography: references.bib  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = TRUE
)

# Utils :
library(here)
library(lubridate)

# Plots : 
library(plotly) # animate / interactive plotting
library(ggplot2) # plotting library
library(gganimate)

# Data : 
library(reshape2) # data shaping
library(dplyr) # data wrangler
library(tidyr) # data manipulation
library(tidyverse)  # data manipulation

```

## Introduction :

This markdown document has been produced with Rstudio. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

The data used for visualization can be downloaded at <https://ourworldindata.org/covid-vaccinations>. It consists of, for every country, and at several dates, the percentage of the population fully or partially vaccinated against COVID19. 

The various sources of 'ourworldindata' are listed and discussed here : 
<https://ourworldindata.org/coronavirus-source-data>

The data columns used here will be : 

_location_: name of the country (or region within a country).
_iso_code_: ISO 3166-1 alpha-3 – three-letter country codes.
_date_: date of the observation.
_people_vaccinated_: total number of people who received at least one vaccine dose. If a person receives the first dose of a 2-dose vaccine, this metric goes up by 1. If they receive the second dose, the metric stays the same.
_people_fully_vaccinated_: total number of people who received all doses prescribed by the vaccination protocol. If a person receives the first dose of a 2-dose vaccine, this metric stays the same. If they receive the second dose, the metric goes up by 1.
_total_boosters_ : Total number of COVID-19 vaccination booster doses administered (doses administered beyond the number prescribed by the vaccination protocol)

*BEWARE :* The data is pretty big (around 29Mb). Careful when charging it whole in your RAM. Cote d'Ivoire replaced by "Cote Ivoire" for easy redeability !

```{r loading and treating data, echo=TRUE}

# columns used : 
MyColumns <- c('location','iso_code','date','people_vaccinated','people_fully_vaccinated','total_boosters','population')
# Get file path : 
filepath <- file.path(here(),'COVID19-Vaccinations','data','owid-covid-data.csv')

# Load whole dataset : 
covid_dataset <- read.table(filepath, 
                  sep=",",
                  header=TRUE,
                  dec=".",
                  row.names=NULL)

# Subset column names we want :
vaccines_dataset <- subset(covid_dataset,select=MyColumns)

# Format all dates : 
vaccines_dataset$formatted_dates <- ymd(vaccines_dataset$date)

# Transform columns in percentage : 
vaccines_dataset$vaccinated <- vaccines_dataset$people_vaccinated / vaccines_dataset$population * 100
vaccines_dataset$fully_vaccinated <- vaccines_dataset$people_fully_vaccinated / vaccines_dataset$population * 100
vaccines_dataset$unvaccinated <- 100 - vaccines_dataset$vaccinated 
vaccines_dataset$semi_vaccinated <- vaccines_dataset$vaccinated -  vaccines_dataset$fully_vaccinated


# Which countries : 
European_Countries <- c('Austria'	,'Belgium','Bulgaria','Croatia'	,
                        'Cyprus'	,'Czechia'	,'Denmark'	,
                        'Estonia'	,'Finland'	,'France'	,
                        'Germany'	,'Greece'	,'Hungary'	,
                        'Ireland','Italy',
                        'Latvia','Lithuania','Luxembourg',
                        'Malta','Netherlands','Poland',
                        'Portugal','Romania','Slovakia',
                        'Slovenia','Spain','Sweden'
                        )
Eu_vaccines_dataset <- subset(vaccines_dataset,location %in% European_Countries )

# Group per vaccination level (semi - full)
Melted_vaccines_dataset <- melt(Eu_vaccines_dataset,
                     measure.vars=c('semi_vaccinated',
                                    'fully_vaccinated'),
                     variable.name = "vaccination_level",
                     value.name = "value"
                     )

test_dataset <- subset(vaccines_dataset,location %in% European_Countries )


start <-  ymd("2021-01-01")
stop <- ymd("2021-08-23")

ToPlot <- subset(Melted_vaccines_dataset,
                 formatted_dates>start & formatted_dates <stop )

```

```{r frozen plot, echo=FALSE, fig.height=5, fig.width=12}

Test <- subset(ToPlot, date == "2021-08-01")
ggplot(data=Test,aes(reorder(value),y=location,
                 fill=vaccination_level,
                 x=value,label=round(value,digits=2)
             )
             )+
        geom_vline(xintercept = 100,color='black')+
        geom_vline(xintercept = 75,color='black')+
        geom_vline(xintercept = 50,color='black')+
        geom_vline(xintercept = 25,color='black')+  
        geom_bar(alpha=0.8,
             position="stack", 
             stat="identity")+
        geom_text(size=2,position = position_stack() )+
        labs(title = 'The european vaccine race | Date: 2021-08-01')+
        theme_classic()+
        theme(legend.position='bottom') +
        scale_x_continuous(limits=c(0,100),expand = c(0, 0))+
        labs(y='Countries',x='Percentage (%)')+
        theme(axis.line = element_blank(),
              axis.ticks.y = element_blank(),
              axis.text.x = element_text(size = 14),
              plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")
              )+
        scale_fill_manual(name='Vaccination level',
                          labels = c("Incomplete vaccination", 
                                      "Fully vaccinated"), 
                          values = c("#34eb8c", "#00AA00"))

```


```{r animated data, echo=FALSE, fig.height=30, fig.width=30}

pp <- ggplot(data=ToPlot,
             aes(y=location,
                 fill=vaccination_level,
                 x=value
                 #label=paste0(round(value,digits=2),'%') 
                 )
             )+
        geom_vline(xintercept = 100,color='grey')+
        geom_vline(xintercept = 50,color='grey',linetype='dashed')+
        geom_vline(xintercept = 70,color='red')+
        geom_vline(xintercept = 0,color='grey')+
        #geom_text(size=4,position = position_stack() )+
        #geom_rect(xmin=60,xmax=70,ymin='Austria',ymax='Sweden',color='grey',alpha=0.2)+
        geom_bar(alpha=0.8,
             position="stack", 
             stat="identity")+
        transition_states(date,
                          transition_length = 2, 
                          state_length = 2)+
        enter_grow()+
        enter_fade()+
        labs(title = 'The european vaccine race | Date: {closest_state}')+
        theme_classic()+
        theme(legend.position='bottom') +
        scale_x_continuous(limits=c(0,100))+
        labs(y='Countries',x='Percentage (%)')+
        shadow_mark() +
        theme(axis.line = element_blank(),
              axis.ticks.y = element_blank(),
              axis.text.x = element_text(size = 14),
              plot.margin = unit(c(0.5,0.5,0.5,0.5), "cm")
              )+
        scale_fill_manual(name='Vaccination level',
                          labels = c("Incomplete vaccination", 
                                      "Fully vaccinated"), 
                          values = c("#a3e6a3", "#00AA00"))

animate(pp, 
        nframes = 570, 
        duration = 25, 
        width = 500, 
        height = 500,
        end_pause = 80
        )

```

