---
title: "WorldAgriculture"
author: "Sebastien Lacube - 2021"
date: "31/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = TRUE
)

# Utils :
library(here)
library(lubridate)

# Plots : 
library(plotly) # animate / interactive plotting
library(ggplot2) # plotting library
library(gganimate)

# Data : 
library(reshape2) # data shaping
library(dplyr) # data wrangler
library(tidyr) # data manipulation
library(tidyverse)  # data manipulation

```

## Introduction :

This markdown document has been produced with Rstudio. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

The data used for visualization can be downloaded at <https://ourworldindata.org/farm-size>. It consists of, for every country, and at several dates, number of farms. Most of the data is quite old (from 1990 to 2000) because there hasn't been any estimate since.  

```{r loading and treating data, echo=TRUE}

# Get file path : 
filepath <- file.path(here(),'WorldAgriculture','data','number-of-farms.csv')


# Load whole dataset : 
farms_dataset <- read.table(filepath, 
                  sep=",",
                  header=TRUE,
                  dec=".",
                  row.names=NULL)

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
