---
title: "Tidy Tuesday - 2021 - Week 36"
author: "Sebastien Lacube"
date: "03/09/2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = TRUE
)

# Utils :
library(tidytuesdayR)

# Plots : 
library(ggplot2) # plotting library

# Data : 
library(tidyverse)  # data manipulation


```

## TidyTuesday - 2021 - Week 36 : Bird Baths

This markdown document has been produced with Rstudio. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

The data used for visualization can be downloaded at <https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-08-31/readme.md>. It consists of data gathered from the original article : <https://doi.org/10.1371/journal.pone.0150899>. 

Variables are as follow : 
- survey_year //	double	// Year of survey
- urban_rural //	character	// Urban or rural location
- bioregions //	character	// Region of Australia
- bird_type //	character	// Bird species
- bird_count //	double	


```{r load data }

# Get data from tidytuesday : https://github.com/rstudio/cheatsheets/blob/master/data-transformation.pdf
tuesdata <- tidytuesdayR::tt_load(2021, week = 36)
bird_baths <- tuesdata$bird_baths


# Clustering birds in each regions by attributes : 
# - Mean of baths between both year
# - Difference between 2014 and 2015 areas
BirdLoss <- bird_baths %>% 
  group_by(bioregions, bird_type,urban_rural,survey_year)%>%
  # Sum similar observations
  summarise(total_bird_count = sum(bird_count)) %>% 
  # Compute differences between 2014 and 2015 measurements
  summarise(diff_total_bird_count = diff(total_bird_count)) %>% 
  # Arrange by difference : 
  arrange(diff_total_bird_count) %>%
  # Overrride grouping : 
  group_by(bioregions,urban_rural) %>%
  # Filter only Pigeons
  filter(grepl('Pigeon',bird_type))



```

## Including Plots

```{r Plotting, echo=FALSE, fig.height=8, fig.width=13, paged.print=TRUE}


# Color theme : 
textcolor = '#ABDBFF'
importanttxtcolor = 'white'
backgroundcolor = 'black'
intermediatecolor = '#FFFFF0'
scalecolor = c("#A40000", "#00aa00")

ggplot(data=BirdLoss,aes(x=diff_total_bird_count,y=bird_type,color=diff_total_bird_count))+
  # Main aesthetics :
  geom_vline(xintercept = 0, color='grey') +
  geom_point(size=3)+
  geom_segment(aes(x=0,y=bird_type,xend=diff_total_bird_count-0.25,yend=bird_type),size=1.5)+
  facet_grid(urban_rural~bioregions)+
  scale_colour_gradient2(low="red", high="blue",mid=intermediatecolor,midpoint = 0,name=' ')+
  # Labs & title : 
  ggtitle('Variations in Pigeon species bathing occurence between seasons \nby bioregion and area type (rural/urban) in Australia.')+
  labs(x='Differences in absolute occurence (winter - summer)',y='Pigeon species')+
  #Theme : 
  theme(
    # Axis : 
    axis.text=element_text(size=12,color=textcolor),
    axis.title.y = element_text(size=16,margin = margin(t = 0, r = 20, b = 0, l = 0),color=importanttxtcolor),
    axis.title.x = element_text(size=16,margin = margin(t = 20, r = 0, b = 0, l = 0),color=importanttxtcolor),
    # Strips : 
    strip.text = element_text(size = 12, colour = textcolor),
    strip.background =element_blank(),
    # Plot : 
    plot.background = element_rect(fill = backgroundcolor, colour = backgroundcolor),
    plot.title = element_text(size=16,hjust = 0.5,margin = margin(t = 0, r = 0, b = 20, l = 0),color=importanttxtcolor),
    plot.caption = element_text(hjust = -3),
    # Legend : 
    legend.position="bottom",
    legend.background = element_rect(fill = backgroundcolor, colour = backgroundcolor),
    legend.key = element_rect(fill = backgroundcolor, colour = backgroundcolor),
    legend.text = element_text(size = 12, colour = textcolor),
    legend.title = element_text(size = 12, colour = textcolor,vjust=1),
    # Panel : 
    panel.spacing = unit(1, "lines"),
    panel.background = element_rect(fill = backgroundcolor, colour = intermediatecolor),    
    panel.grid.major.x =element_blank(),
    panel.grid.major.y =element_blank(),
    panel.grid.minor.x =element_blank(),
    panel.grid.minor.y =element_blank()
  )




```

```{r No difference, fig.height=4, fig.width=10}


# Clustering birds in each regions by attributes : 
# - Mean of baths between both year
# - Difference between rural and urban areas
NumberOfDoves <- bird_baths %>% 
  group_by(bioregions, bird_type,urban_rural,survey_year)%>%
  # Sum similar observations
  summarise(total_bird_count = sum(bird_count)) %>% 
  # Split survey_year column for plot : 
  spread(survey_year, total_bird_count) %>% 
  # Overrride grouping : 
  group_by(bioregions,urban_rural) %>%
  # Remove NA rows :
  filter(bioregions != 'NA')   %>%
  # Remove last NA column (global sum):
  select(-6)


colnames(NumberOfDoves) <- c(colnames(NumberOfDoves[1:3]),'Winter','Summer')

NumberOfDoves$Total <- NumberOfDoves$Winter + NumberOfDoves$Summer

ToPlot <- NumberOfDoves  %>%
  group_by(urban_rural,bioregions) %>%
  slice_max(Summer,n=5,with_ties = FALSE)
  
ggplot(data=ToPlot,aes(x=Winter,y=bird_type,yend=bird_type,xend=Summer,color=urban_rural))+
  ggtitle('Variations in Dove species bathing occurence between summer and winter \nby bioregion in Australia and Area type (rural/urban).')+
  geom_vline(xintercept = 0, color='black') +
  geom_point(aes(x=Winter),size=2)+
  geom_point(aes(x=Summer),size=2)+
  geom_segment(arrow = arrow(length=unit(0.1,"cm"), ends="first", type = "closed"),color='black')+
  facet_grid(urban_rural~bioregions)+
  labs(x='Difference in absolute occurence (winter - summer)',y='Dove species')+
  theme_bw()

```

