---
title: "Tidy Tuesday - 2021 - Week 37 - Formula One"
author: "Sebastien Lacube"
date: "03/09/2021"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = TRUE
)

# Utils :
library(tidytuesdayR)

# Plots : 
library(ggplot2) # plotting library
library(ggExtra) # extras for platting library

# Data : 
library(tidyverse)  # data manipulation
library(afex) # to run anova

```

## TidyTuesday - 2021 - Week 37 : Formula One

This markdown document has been produced with Rstudio. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

The data used for visualization can be downloaded at <https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-08-31/readme.md>. The data this week comes from the Ergast API, which has a CC-BY license. H/t to Sara Stoudt for sharing the link to the data by way of Data is Plural! 

Variables from the API include several tables : 
- circuits 
- drivers & driver standings
- races from whole year of Formula One racing

I personally added one more dataset to do the link between nationalities and country names : <https://raw.githubusercontent.com/knowitall/chunkedextractor/master/src/main/resources/edu/knowitall/chunkedextractor/demonyms.csv>


```{r load and manipulate data }

# Find correspondance between nationality & country : 
NationalityCountry <- readr::read_csv('https://raw.githubusercontent.com/knowitall/chunkedextractor/master/src/main/resources/edu/knowitall/chunkedextractor/demonyms.csv',col_names =FALSE) 
colnames(NationalityCountry) <- c('nationality','country') 
AllNationalityCountry <- NationalityCountry %>%
                            # Add the right nationalities : 
                            add_row(nationality='Monegasque',country='Monaco')%>%
                            add_row(nationality='East German',country='East Germany')
  
# Load circuit data : 
Circuits_by_country <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/circuits.csv') %>% 
                            rename(circuit.name=name) %>% 
                            rename(circuit.location=location) %>% 
                            rename(circuit.country=country) %>%
                            select(circuitId,circuit.name,circuit.country) %>%
                            # Properly recode countries : 
                            mutate(circuit.country = fct_recode(circuit.country, "United States" ="USA")) %>%
                            mutate(circuit.country = fct_recode(circuit.country, "United Kingdom" ="UK"))



# Drivers data & standings : 
drivers_by_country <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/drivers.csv') %>% 
                          left_join(AllNationalityCountry) %>%
                          rename(driver.country=country) %>%
                          unite(driver.name, c("forename", "surname"),sep = ' ') %>%
                          select(driverId,driver.name,driver.country) 

constructors_by_country <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/constructors.csv') %>% 
                          left_join(AllNationalityCountry) %>%
                          rename(constructor.country=country) %>%
                          rename(constructor.name=name) %>%
                          select(constructorId,constructor.name,constructor.country) 

# Races data : 
races_by_country <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/races.csv') %>% 
    inner_join(Circuits_by_country,by=c('circuitId')) %>%
    rename(race.name=name) %>%
    rename(race.year=year) %>%
    rename(race.round=round) %>%
    select(raceId,race.name,circuit.name,circuit.country,race.year,race.round) 

# Driver standings : 
driver_standings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/driver_standings.csv') %>%
  left_join(races_by_country,by=c('raceId')) %>%
  inner_join(drivers_by_country,by=c('driverId')) %>%
  distinct_all() %>%
  select(race.year,race.round,race.name,circuit.name,driver.name,points,position,wins,circuit.country,driver.country)%>%
  mutate(IsHome =  (circuit.country == driver.country ) )



# Others data : 
# lap_times <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/lap_times.csv')
# pit_stops <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/pit_stops.csv')
# qualifying <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/qualifying.csv')
# results <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/results.csv')
# seasons <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/seasons.csv')
# status <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/status.csv')


```

```{r drivers data analysis, fig.height=4.5, fig.width=6}
# 

ToPLot <- subset(driver_standings,!is.na(IsHome)) %>%
  mutate(IsHomeChar = as.character(IsHome)) %>%
  mutate(race.type = fct_recode(IsHomeChar, "Home" ="TRUE","Outside" ="FALSE"))

one.way <- aov( position ~ race.type, data = ToPLot)
summary(one.way)

two.way <- aov( position ~ race.type + driver.country, data = ToPLot)
summary(two.way)

Summary <- subset(driver_standings,!is.na(IsHome)) %>%
  mutate(IsHomeChar = as.character(IsHome)) %>%
  mutate(race.type = fct_recode(IsHomeChar, "Home" ="TRUE","Outside" ="FALSE")) %>%
  group_by(IsHome) %>%
  summarise(mean=mean(position), sd=sd(position))

```


```{r constructor data analysis }


# Driver standings : 
constructor_standings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-07/constructor_standings.csv') %>%
  left_join(races_by_country,by=c('raceId')) %>%
  inner_join(constructors_by_country,by=c('constructorId')) %>%
  distinct_all() %>%
  select(race.year,race.round,race.name,circuit.name,constructor.name,points,position,wins,circuit.country,constructor.country)%>%
  mutate(IsHome =  (circuit.country == constructor.country ) )


Constructors_data <- subset(constructor_standings,!is.na(IsHome)) %>%
  mutate(IsHomeChar = as.character(IsHome)) %>%
  mutate(race.type = fct_recode(IsHomeChar, "Home" ="TRUE","Outside" ="FALSE")) 

Constructors_by_racetype <- Constructors_data%>%
  group_by(race.type) %>%
  summarise(mean=mean(position), sd=sd(position),n=n())


one.way <- aov( position ~ race.type, data = Constructors_data)
summary(one.way)

```

## Data manipulation

```{r driver_performances analysis, fig.height=8, fig.width=8}

MyDrivers <- c('Lewis Hamilton','Fernando Alonso','Sebastian Vettel','Pierre Gasly','Sergio Pérez','Lance Stroll','George Russell','Lando Norris','Esteban Ocon','Antonio Giovinazzi','Carlos Sainz')

Test <- subset(driver_standings,!is.na(IsHome)) %>%
  mutate(IsHomeChar = as.character(IsHome)) %>%
  mutate(race.type = fct_recode(IsHomeChar, "Home" ="TRUE","Outside" ="FALSE")) %>%
  group_by(IsHome) %>%
  filter(driver.name %in% MyDrivers)%>%
  group_by(race.type,driver.name,driver.country) %>%
  summarise(mean=mean(position), sd=sd(position),n=n())

Pivoted_drivers_data <- Test %>% 
  pivot_wider( id_cols = driver.name, names_from = race.type, values_from = mean:n)%>%
  arrange(mean_Outside) %>%
  mutate(driver.name=factor(driver.name, levels=driver.name))

Pivoted_drivers_data$driver.name = with(Pivoted_drivers_data, reorder(driver.name, mean_Outside))


# Theme : 
textcolor = '#E7E6E6'
importanttxtcolor = '#E7E6E6'
backgroundcolor = 'black'
intermediatecolor = '#E7E6E6'

myplot <- ggplot(data=Pivoted_drivers_data)+
  # Main aesthetics :
  # geom_point(aes(x=mean_Home,y=driver.name),
  #            color='#FF2800',
  #            size=1.5)+
  # geom_point(aes(x=mean_Outside,y=driver.name),
  #            color='#FFC000',
  #            size=1.5)+
  geom_segment(aes(x=mean_Home,y=driver.name,xend=mean_Outside,yend=driver.name,color=mean_Outside-mean_Home),
               size=1.5,
               arrow = arrow(length = unit(0.2, "cm")),
               lineend = 'round',
               linejoin = 'round'
               )+
  geom_line(aes(x=mean_Home,y=driver.name,group=mean_Outside),
             color='#FF2800',
             size=3,
            )+
  labs(x='Average position',y='')+
  #scale_y_discrete(position = "right")+
  scale_colour_gradient2(low="#FFC000", high="#FF2800",mid='#4A595F',midpoint = 0,breaks=c(-2,2),labels=c('Better\n Outside','Better\n Home'),name='Preferences')+
  theme(
    # Axis : 
    axis.text=element_text(size=12,color=textcolor),
    axis.text.x = element_text(size=16,color=textcolor),
    axis.text.y = element_text(size=16,color=textcolor),
    axis.title.x = element_text(size=16,color=importanttxtcolor),
    # Strips : 
    strip.text = element_text(size = 12, colour = textcolor),
    strip.background =element_blank(),
    # Plot : 
    plot.background = element_blank(),
    plot.title = element_text(size=16,hjust = 0.5,margin = margin(t = 0, r = 0, b = 20, l = 0),color=importanttxtcolor),
    plot.caption = element_text(hjust = -3),
    # Legend : 
    legend.position="bottom",
    legend.background = element_blank(),
    legend.key = element_rect(fill = backgroundcolor, colour = backgroundcolor),
    legend.text = element_text(size = 12, colour = textcolor),
    legend.title = element_text(size = 16, colour = textcolor,vjust=1),
    # Panel : 
    panel.spacing = unit(1, "lines"),
    panel.background = element_rect(fill = backgroundcolor, colour = backgroundcolor),
    panel.grid.major.x =element_blank(),
    panel.grid.major.y =element_blank(),
    panel.grid.minor.x =element_blank(),
    panel.grid.minor.y =element_blank()
  )

ggsave(plot = myplot, file = "graph2.png", 
       type = "cairo-png",  bg = "transparent",
       width = 20, height = 14, units = "cm", dpi = 800)

```


```{r Summarize drivers data (n), echo=FALSE, fig.height=8, fig.width=13, paged.print=TRUE}

subset(driver_standings,!is.na(IsHome)) %>%
  mutate(IsHomeChar = as.character(IsHome)) %>%
  mutate(race.type = fct_recode(IsHomeChar, "Home" ="TRUE","Outside" ="FALSE")) %>%
  group_by(driver.name,IsHome) %>%
  summarise(mean=mean(position), sd=sd(position),n=n())%>%
  filter(driver.name %in% MyDrivers)

```

